---
title: "A brief guide to using the wcfish R package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 1. Installation

Let's begin by loading the wcfish package and its dependencies.

```{r setup, results = "hide"}
# Load package
library(wcfish)
library(tidyverse)
```


## 2. Datasets

The package includes the following West Coast fisheries datasets (listed below):

__Coastwide datasets__

- West Coast commercial fisheries species: `?taxa`
- West Coast commercial fishing ports: `?ports`
- West Coast commercial fishing blocks shapefile: `?blocks`

__California (CDFW) datasets__

- Annual commercial landings by source and species: `?cdfw_waters`
- Annual commercial landings by port and species: `?cdfw_ports`
- Annual number of licensed commercial fishers by area of residence: `?cdfw_n_comm_fishers`
- Annual number of registered commercial fishing vessels by length class: `?cdfw_n_comm_vessels_length`
- Annual number of registered commercial fishing vessels by port complex: `?cdfw_n_comm_vessels_port`
- Annual number of registered commercial fishing vessels: `?cdfw_n_comm_vessels`
- Annual CPFV landings by port complex and species: `?cdfw_cpfv`
- Annual CPFV participation and effort by port complex and species: `?cdfw_cpfv_effort`
- Annual kelp harvest by bed type: `?cdfw_kelp`
- Annual kelp harvest overall: `?cdfw_kelp_tots`

__PACFIN datasets__

- Species meta-data: `?pacfin_species`
- Ports and port complexes meta-data: `?pacfin_ports`
- Annual commercial landings by state (CA/OR/WA), port complex, and species: `?pacfin_all6`
- Monthly commercial Dungeness crab landings by state (CA/OR/WA) and port complex: `?pacfin_crab2`

__Other West Coast fisheries datasets__

- Monthly commercial landings in CA by port complex and species from the SWFCC: `?swfsc`
- Annual commercial landings in CA by port complex, gear type, and species from the CALCOM: `?calcom`


## 3. Inspecting datasets

Every dataset is documented with metadata. To view the metadata for a dataset, use the `?` command. For example, if you wanted to look at the meta-data for the `cdfw_ports` dataset, you would do the following:

```{r }
# Examine the metadata for the cdfw_port dataset
?cdfw_ports
```

This will pop up a description of the dataset and its columns in the "Help" screen of your R text editor.


## 4. Saving the dataset as an object

You can save the dataset as an object in your R environment using the assingment operator:

```{r }
# Save a dataset in your R environment
data <- wcfish::cdfw_ports
```


## 5. Preparing a quick plot of the dataset

We encourage you to work with the data and plot it based on your own needs. However, we also provide a `plot_data` function for generating a quick plot of the dataset of interest. Note that this only works for the CDFW datasets.

```{r fig.width=7, fig.height=4}
# Quickly plot a CDFW dataset
wcfish::plot_data("cdfw_ports")
```











